---
title: "simpleRegress"
author: "Joel Calderin"
date: "Fall 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 


## Model LENGTH ~ AGE


```{r}
AGE <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(AGE)
summary.lm(AGE)
```



## Model LENGTH ~ HEADLEN


```{r}
HEADLEN <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(HEADLEN)
summary.lm(HEADLEN)
```


## Model LENGTH ~ CHEST
   
```{r}
CHEST <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(CHEST)
summary.lm(CHEST)
```

## Model LENGTH ~ HEADWTH


```{r}
HEADWTH <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(HEADWTH)
summary.lm(HEADWTH)
```

## Model LENGTH ~ WEIGHT


```{r}
WEIGHT <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(WEIGHT)
summary.lm(WEIGHT)
```


## Discussion and Conclusion 

The model that showed the largest adjusted R-Squared is LENGTH ~ HEADLEN at 0.8434. This was the best model to determine the bear's length. The lowest adjusted R-Squared was model LENGTH ~ AGE at 0.5073. This model would be the worst one to determine the bear's length. 




